---
title: "Class 11: Structural Bioinformatics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## The PDB database for biomolecular structure data

```{r}
data <- read.csv("Data Export Summary.csv")
data
```
Q1: 
determine the percentage of structures solved by X-Ray and Electron Microscopy. 
```{r}
(data$Total/sum(data$Total)) * 100
```



determine what proportion of structures are protein?

```{r}
(sum(data$Proteins)/sum(data$Total)) *100
```



```{r}
data <- read.csv("Data Export Summary.csv")
data
```


Total number of entries
```{r}
sum(data$Total)
```


proportion of entries from each method
```{r}
round((data$Total/sum(data$Total)) * 100, 2)
```

Proportion that are protein
```{r}
round((sum(data$Proteins) / sum(data$Total)) * 100, 2)
```


## HIV-Pr

```{r}
library(bio3d)
pdb <- read.pdb("1hsg")
## Note: Accessing on-line PDB file
print(pdb)
```

```{r}
read.pdb("1HSG.pdb.txt")
```

Q6. How many amino acid residues are there in this pdb object and what are the two nonprotein residues?
198 residues (CA)
water and MK1 are the two nonprotein residues

```{r}
ligand <- atom.select(pdb, "ligand", value = TRUE)
write.pdb(ligand, file="1hsg_ligand.pdb")
```







