---
title: "class 09"
author: "Chris"
date: "10/29/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
wisc.df <- read.csv("WisconsinCancer.csv")

head(wisc.df)
```


we have `r nrow(wisc.df)` samples 
```{r}
nrow(wisc.df)
```



How many benign vs malignant patients?
```{r}
table(wisc.df$diagnosis)
```


Rescale matrix to not include unnecessary data
```{r}
wisc.data <- as.matrix(wisc.df[,3:32])

row.names(wisc.data) <- wisc.df$id

head(wisc.data)
```

store diagnosis for reference in future as separate vector

```{r}
diagnosis <- wisc.df$diagnosis
```



Q1. How many observations are in this dataset?

```{r}
nrow(wisc.df)
```

Q2. How many of the observations have a malignant diagnosis?

```{r}
table(wisc.df$diagnosis)
```

Q3. How many variables/features in the data are suffixed with _mean?

```{r}
grep( "_mean", colnames(wisc.df), value = TRUE)
```

The functions dim(), nrow(), table(), length() and grep() may be useful for answering the first 3 questions above.


```{r}
length(grep( "_mean", colnames(wisc.df)))
```


2. Principal Component Analysis
Performing PCA
The next step in your analysis is to perform principal component analysis (PCA) on wisc.data.

It is important to check if the data need to be scaled before performing PCA. Recall two common reasons for scaling data include:

The input variables use different units of measurement.
The input variables have significantly different variances.
Check the mean and standard deviation of the features (i.e. columns) of the wisc.data to determine if the data should be scaled. Use the colMeans() and apply() functions like you’ve done before.

means are highly variable
```{r}
round(colMeans(wisc.data),3)


```

so are the standard deviation, so scale is needed
```{r}
round(apply(wisc.data, 2 ,sd),3)
```

set scale to true to scale data
```{r}
wisc.pr <- prcomp(wisc.data, scale. = TRUE)

summary(wisc.pr)
```





Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?
.4427 (proportion of variance)

Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
3 (cumulative proportion)

Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?
7



```{r}
biplot(wisc.pr)
```


```{r}
plot( wisc.pr$x[,1], wisc.pr$x[,2], col = diagnosis, xlab = "PC1", ylab = "PC2")
```



```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,3], col = diagnosis, xlab = "PC1", ylab = "PC3")
```

variance equation
```{r}
pr.var <- wisc.pr$sdev^2
head(pr.var)
```
Scree plot
```{r}
pve <- pr.var/sum(pr.var)

plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```


```{r}
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```


```{r}
## ggplot based graph
#install.packages("factoextra")
library(factoextra)
fviz_eig(wisc.pr, addlabels = TRUE)
```

scale data
```{r}
data.scaled <- scale(wisc.data)
```
```{r}
data.dist <- dist(data.scaled, method = "euclidean")

wisc.hclust <- hclust(data.dist,method = "complete" )
```

make dendrogram
```{r}
plot(wisc.hclust)
abline(h = 19, col = "red", lty = 2)

wisc.hclust.clusters <- cutree(wisc.hclust, k = 4)
```



```{r}
table(wisc.hclust.clusters, diagnosis)
```





```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[, 1:7]), method ="ward.D2")
plot(wisc.pr.hclust)
```

```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```

Cluster 1 most likely malignant, cluster 2 most likely benign
28 false positives, 24 false negatives
```{r}
table(grps, diagnosis)
```


```{r}
plot(wisc.pr$x[,1:2], col=grps)
```


```{r}
plot(wisc.pr$x[,1:2], col=diagnosis)

```

```{r}
g <- as.factor(grps)
levels(g)

g <- relevel(g,2)
levels(g)

plot(wisc.pr$x[,1:2], col=g)
```






```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```


```{r}
plot(wisc.pr$x[,1:2], col=g)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")

```

Q17. Which of these new patients should we prioritize for follow up based on your results?
Patient 2


