---
title: "Class 12: Structural Bioinformatics II"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prepare for docking

We want to produce a protein-only PDB file and a drug only PDB file.
```{r}
library(bio3d)
file.name <- get.pdb("1hsg")
```


```{r}
hiv <- read.pdb("1hsg.pdb")
hiv
```


Q1: What is the name of the two non protein resid values in this structure? What does resid
correspond to and how would you get a listing of all reside values in this structure? 
Water and MK1


```{r}
pdb <- read.pdb("1hsg.pdb")
protein <- atom.select(pdb, "protein", value = TRUE)
write.pdb(protein, file= "1hsg_protein.pdb")
protein
```

```{r}
ligand <- atom.select(pdb, "ligand", value = TRUE)
write.pdb(ligand, file = "1hsg_ligand.pdb")
ligand
```


Q2: Can you locate the binding site visually? Note that
crystal structures normally lack hydrogen atoms, why? 
The empty area inside the molecule, too small?

Q3: Look at the charges. Does it make sense (e.g. based on your knowledge of the
physiochemical properties of amino acids)?




##Process the docking results
```{r}
library(bio3d)
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
```



```{r}
# res <- read.pdb("all.pdbqt", multi=TRUE)
ori <- read.pdb("ligand.pdbqt")
rmsd(ori, res)

```
















