---
title: "Class 7 R Functions and Packagaes"
author: "Chris"
date: "10/22/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R functions revisited

Source my functions from last day

```{r}
source("http://tinyurl.com/rescale-R")
```

```{r}
rescale(1:10)
```


```{r}
rescale(c(1, 10, 5, NA, 6))
```

```{r}
rescale2(1:10)
```



Write a new function called bothna() to find where there are NA elements in two input vectors.


First make some simple input where I know the answer
```{r}
x <- c( 1, 2, NA, 3, NA)

y <- c(NA, 3, NA, 3, 4)
```

Looked online and found **is.na()** function
```{r}
is.na(x)
```

and the **which()** function tells me where the TRUE values are
```{r}
which(is.na(x))
```

```{r}
which(is.na(y))
```

```{r}
is.na(x)
is.na(y)
```

The AND function requires two input TRUE to give output
```{r}
is.na(x) & is.na(y)
```

```{r}
sum(is.na(x) & is.na(y))
```

Taking the **sum()** of TRUE FALSE vector will tell me how many TRUE elements I have. This is my working snippet!!
```{r}
sum(c(TRUE, TRUE, FALSE, TRUE))
```


Now turn it into a function
```{r}
both_na <- function(x, y) {
  sum(is.na(x) & is.na(y))
}
```



```{r}
both_na(x, y)
```

Test function
```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
```

```{r}
both_na(x, y1)
```

```{r}
both_na(x, y2)
```

```{r}
x2 <- c(NA, NA)
```
```{r}
both_na(x2, y2)
```


```{r}
plot(1:10, col = c("red", "blue"))
```

repeats vector x until same length pattern 1, NA, NA
```{r}
x <- c(1, NA, NA)
y3 <- c(1, NA, NA, NA, NA, NA, NA)
both_na(x, y3)
```


```{r}
length(x)
length(y3)
```






```{r}
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
```
```{r}
mean(student1[-which.min(student1)])
mean(student2[-which.min(student2)], na.rm = TRUE)
```


We have our working code now turn it into a first function

```{r}
grade <- function (x) {
  if(any(is.na(x)) == TRUE) {
  warning("Student is missing a homework")  
  }
  mean(x[-which.min(x)], na.rm = TRUE)
}
```

```{r}
grade(student1)
grade(student2)
```

```{r}
grade2 <- function (x) {
  if(any(is.na(x)) == TRUE) {
  warning("Student is missing a homework")  
  }
  (sum(x[-which.min(x)], na.rm = TRUE))/(length(x)-1)
}
```

```{r}
student3 <- c(100, NA, NA, NA, NA, NA, NA, NA)
grade2(student3)
```

```{r}
url <- "https://tinyurl.com/gradeinput"
hw <- read.csv(url, row.names = 1)
```

```{r}
apply(hw, 1, grade)
```










